---
title: "Lab 4"
author: "Maureen McDonald and Omima Hassan"
date: "10/17/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(tidyverse)
library(plyr)
```

## **Loading the Data: Cyclone Football**
```{r}
library(readxl)
defense <- read_excel('cyclonesFootball2018.xlsx', sheet='Defensive')
offense <- read_excel('cyclonesFootball2018.xlsx', sheet='Offensive')
bio <- read_excel('cyclonesFootball2018.xlsx', sheet='Biography')
```

## **Part One: Data Cleaning**

#### Changing the player and opponent names to factors.
```{r}
DefName= defense$Name= as.factor(defense$Name)
DeffOpponent_Opponent= defense$Opponent_Opponent= as.factor(defense$Opponent_Opponent)

OffName= offense$Name=offense$Name= as.factor(offense$Name)
OffOpponent_Opponent=offense$Opponent_Opponent= as.factor(offense$Opponent_Opponent)

Name= bio$Name= as.factor(bio$Name)
```


#### Changing all offensive and defensive statistics as well as weight to numeric values.
```{r}
Tackles_Solo= defense$Tackles_Solo=as.numeric(defense$Tackles_Solo)

Tackles_ASST= defense$Tackles_ASST=as.numeric(defense$Tackles_ASST)

Tackles_TFL= defense$Tackles_TFL=as.numeric(defense$Tackles_TFL)

Tackles_Sack= defense$Tackles_Sack=as.numeric(defense$Tackles_Sack)

Turnover_FF= defense$Turnover_FF=as.numeric(defense$Turnover_FF)

Turnover_FR= defense$Turnover_FR=as.numeric(defense$Turnover_FR)

Turnover_INT= defense$Turnover_INT=as.numeric(defense$Turnover_INT)

Pass_QBH= defense$Pass_QBH= as.numeric(defense$Pass_QBH)

Pass_PB= defense$Pass_PB= as.numeric(defense$Pass_PB)

Receiving_REC= offense$Receiving_REC= as.numeric(offense$Receiving_REC)

Receiveing_YDS= offense$Receiving_YDS=as.numeric(offense$Receiving_YDS)

Receiveing_TD= offense$Receiving_TD=as.numeric(offense$Receiving_TD)

Rushing_ATT= offense$Rushing_ATT=as.numeric(offense$Rushing_ATT)

Rushing_YDS= offense$Rushing_YDS=as.numeric(offense$Rushing_YDS)

Rushing_TD= offense$Rushing_TD=as.numeric(offense$Rushing_TD)

'Passing_CMP-ATT'= offense$'Passing_CMP-ATT'=as.numeric(offense$'Passing_CMP-ATT')

Passing_YDS= offense$Passing_YDS=as.numeric(offense$Passing_YDS)

Passing_TD= offense$Passing_TD=as.numeric(offense$Passing_TD)

Passing_INT= offense$Passing_INT=as.numeric(offense$Passing_INT)

Weight= bio$Weight= as.numeric(bio$Weight)
```

#### Separating the height column into two columns that are each numeric.


```{r}
bio= bio %>% separate(Height, into= c("Feet", "Inches"))

Feet= bio$Feet=as.numeric(bio$Feet)
Inches= bio$Inches=as.numeric(bio$Inches)

FeetAsInches = Feet * 12
newInches = FeetAsInches + Inches
newInches = as.numeric(newInches)

bio = bio %>% mutate(Height_Inches = newInches)
bio = bio %>% select('Name', 'Weight', 'Height_Inches', 'Hometown')
head(bio)
```
* Height is now in inches 

#### Cleaned data frames
```{r}
defClean= defense
offClean= offense
bioClean= bio
```

## **Part Two: Tidying (A)**
```{r}
redefClean= defClean %>% gather(key= Defense_Stat, value = Statistic, Tackles_Solo:Pass_PB)
head(redefClean)
tempData = redefClean %>% filter(Defense_Stat == "Tackles_Solo",Opponent_Opponent == "West Virginia" | Opponent_Opponent == "Kansas State", Statistic != 0)
```


#### Histogram representing each defensive statistic
```{r}
ggplot(redefClean, aes(Statistic))+ geom_histogram(binwidth = 1) + facet_wrap(~Defense_Stat) + labs(title = "Relationship between defensive skills", x= "Type of Defensive Statistic", y = "Amount of Defensive Statistic")
```

#### ISU Defense relationship between West Virginia and Kansas State
```{r}
tempData = tempData %>% spread(Opponent_Opponent, Statistic)

ggplot(tempData, aes(tempData$`West Virginia`, tempData$`Kansas State`)) + geom_point() + labs (title = "ISU Defense against different Schools", x = "West Virginia", y = "Kansas State")
```



## **Part Two: Tidying (B)**

#### Separating Hometown into two columns, City and State
```{r}
bioClean= bioClean%>% separate(Hometown, into = c("City", "State"), sep = ",")
head(bioClean)
```

#### Determining how many players are from each state.
```{r}
count(bioClean$State)
```

## **Part Three: Joining Data Frames**
```{r}
bo= inner_join(offClean, bioClean, by= c("Name"))
```

#### Determining if a relationship exists between the weight of a player and their receiving yards.
```{r}
ggplot(bo, aes(Weight, Receiving_YDS))+ geom_point() + labs(title = "Weight vs. Receiving Yards", x = "Weight", y= "Amount of Receiving Yards")
```

* From the plot it's been determined that there isn't a clear representation to whether or not a relationship exists between weight of a player and receiving yards.


#### Determining whether a relationship exists between the weight of a player and the amount of tackles they have.
```{r}
bd= full_join(defClean, bioClean, by= c("Name"))
head(bd)
ggplot(bd, aes(Weight, Tackles_Solo))+ geom_point() + labs(title = "Weight vs. Tackles", x = "Weight", y= "Amount of Tackles")
```

* From the plot it's been determined that there exists a negative relationship between weight and solo tackles. As your weight increases your solo tackles are decreased.